Listings of equations, figures, tables and sections of the article '{\it \bf Musical elements in the discrete-time representation of sound}' and of the scripts in the MASS toolbox

### 2017-08-01 article/acm/body.tex
finished listing of equations

\section{Characterization of the discrete-time musical note} \label{sec:discNote}\label{sec:notaDisc}
In diverse artistic and theoretical contexts, music is conceived as comprising fundamental units referred to as notes, ``atoms'' that constitute music itself~\cite{Wisnick, Lovelock, Webern}.
These notes are understood as central elements of certain musical paradigms.
In a cognitive perspective, the notes are seen as discretized elements that facilitate and enrich the flow of information through music~\cite{Roederer, Lacerda}.
Canonically, the basic characteristics of a musical note are duration, loudness, pitch and timbre~\cite{Lacerda}, which can be quantified using the evenly time spaced sonic samples.

@@

\noindent where $\zeta$ is the distance between ears $\zeta$, known to be $\zeta \approx 21.5cm$ in an adult human. Straightforward calculations result in ITD:

@@

\section{Variation in the basic musical note}\label{sec:internalVar}\label{sec:varInternas}
The basic digital music note defined in Section~\ref{sec:notaDisc} has the following parameters: duration, pitch, intensity (volume) and timbre. This is a useful and paradigmatic model, but it does not exhaust all the aspects of a musical note. First of all, characteristics of the note change along the note itself~\cite{Chowning}. For example, a $3s$ piano note has intensity with abrupt rise at the beginning and progressive decay, has spectrum variations with harmonics decaying and some others emerging along time. These variations are not mandatory, but they are used in sound synthesis for music because they reflect how sounds appear in nature. This is considered so important that there is a rule of thumb: to make a sound that incites interest by itself, arrange internal variations on it~\cite{Roederer}.
To explore all the ways in which variations occur within a note is out of the scope of any work, given the sensibility of the human ear and the complexity of human sound cognition. In the following, primary resources are presented to produce variations in the basic note. It is worthwhile to recall that all the relations in this and other sections are implemented in Python and published in the public domain \massa\ toolbox. The musical pieces \emph{Transit para meter}, \emph{Vibrates and shakes}, \emph{Tremolos, vibratos and the frequency}, \emph{Little train of impulsive hillbillies}, \emph{Noisy band}, \emph{Bela rugosi}, \emph{Children's chorus}, \emph{ADa and SaRa} were made to validate and illustrate concepts of this section. The code that synthesizes these pieces is also part of the toolbox\cite{MASSA}.

@@

  \item The brown noise received this name after Robert Brown, who described the Brownian movement\footnote{Although its origin is disparate with its color association, this noise became established with this specific name, in musical contexts. Anyway, this association can be considered satisfactory once violet, blue, white and pink noises are more strident and associated with more vivid colors~\cite{Cook,guillaume}.} What characterizes brown noise is the decrease of $6dB$ per octave, with $\alpha_i$ in Equations~\ref{eq:rosa} being:

@@

\newcommand{\OneColEqu}[1]{%
\end{multicols}%
\begin{twocolequfloat}%
\begin{equation}
\{t_i'\}_0^{\Lambda-1} =\{(1+a_i) . t_i\}_0^{\Lambda-1} = \left \{ \left [ 1+M.\sin \left ( f'.2\pi\frac{i}{f_s -1} \right ) \right] .P .\sin \left ( f.2\pi\frac{i}{f_s -1} \right ) \right \}_0^{\Lambda-1} = \\ 
                        =  \left\{P.\sin \left( f.2\pi\frac{i}{f_s -1}  \right ) +  \frac{P.M}{2} \left [ \sin \left( (f-f').2\pi\frac{i}{f_s -1}  \right )  + \sin \left( (f+f').2\pi\frac{i}{f_s -1}  \right ) \right ] \right \}_0^{\Lambda-1}
\end{equation}
\end{twocolequfloat}%
\begin{multicols}{2}%
}

@@

\begin{equation}
P^{ \{ j_k \} }_{ \{ i_{k} \}}
\end{equation}

>>>>

\section{Characterization of the musical note in dicrete-time audio} \label{sec:discNote}\label{sec:notaDisc}
In diverse artistic and theoretical contexts, music is conceived as comprising fundamental units referred to as notes, ``atoms'' that constitute music itself~\cite{Wisnick, Lovelock, Webern}.
These notes are understood as central elements of certain musical paradigms.
In a cognitive perspective, the notes are seen as discretized elements that facilitate and enrich the flow of information through music~\cite{Roederer, Lacerda}.
Canonically, the basic characteristics of a musical note are duration, loudness, pitch and timbre~\cite{Lacerda}, which can be quantified using the evenly time spaced sonic samples.

@@

\noindent where $\zeta$ is the distance between ears $\zeta$, known to be $\zeta \approx 21.5cm$ in an adult human. Straightforward calculations result in Interaural Time Difference:

@@

\section{Variation in the basic note}\label{sec:internalVar}\label{sec:varInternas}
The basic digital music note defined in Section~\ref{sec:notaDisc} has the following parameters: duration, pitch, intensity (volume) and timbre. This is a useful and paradigmatic model, but it does not exhaust all the aspects of a musical note. First of all, characteristics of the note change along the note itself~\cite{Chowning}. For example, a $3s$ piano note has intensity with abrupt rise at the beginning and progressive decay, has spectrum variations with harmonics decaying and some others emerging along time. These variations are not mandatory, but they are used in sound synthesis for music because they reflect how sounds appear in nature. This is considered so important that there is a rule of thumb: to make a sound that incites interest by itself, arrange internal variations on it~\cite{Roederer}.
To explore all the ways in which variations occur within a note is out of the scope of any work, given the sensibility of the human ear and the complexity of human sound cognition. In the following, primary resources are presented to produce variations in the basic note. It is worthwhile to recall that all the relations in this and other sections are implemented in Python and published in the public domain \massa\ toolbox. The musical pieces \emph{Transit para meter}, \emph{Vibrates and shakes}, \emph{Tremolos, vibratos and the frequency}, \emph{Little train of impulsive hillbillies}, \emph{Noisy band}, \emph{Bela rugosi}, \emph{Children's chorus}, \emph{ADa and SaRa} were made to validate and illustrate concepts of this section. The code that synthesizes these pieces is also part of the toolbox\cite{MASSA}.

@@

  \item The brown noise received this name after Robert Brown, who described the Brownian movement\footnote{Although its origin is disparate with its color association, this noise became established with this specific name in musical contexts. Anyway, this association can be considered satisfactory once violet, blue, white and pink noises are more strident and associated with more vivid colors~\cite{Cook,guillaume}.} What characterizes brown noise is the decrease of $6dB$ per octave, with $\alpha_i$ in Equations~\ref{eq:rosa} being:

@@

\begin{equation}\label{eq:specAM}
\begin{split}
\{t_i'\}_0^{\Lambda-1} =\{(1+a_i) . t_i\}_0^{\Lambda-1} = \left \{ \left [ 1+M.\sin \left ( f'.2\pi\frac{i}{f_s -1} \right ) \right] .P .\sin \left ( f.2\pi\frac{i}{f_s -1} \right ) \right \}_0^{\Lambda-1} = \\ 
                        =  \left\{P.\sin \left( f.2\pi\frac{i}{f_s -1}  \right ) +  \frac{P.M}{2} \left [ \sin \left( (f-f').2\pi\frac{i}{f_s -1}  \right )  + \sin \left( (f+f').2\pi\frac{i}{f_s -1}  \right ) \right ] \right \}_0^{\Lambda-1}
\end{split}
\end{equation}

% \newcommand{\OneColEqu}[1]{%
% \end{multicols}%
% \begin{twocolequfloat}%
% \begin{equation}
% \{t_i'\}_0^{\Lambda-1} =\{(1+a_i) . t_i\}_0^{\Lambda-1} = \left \{ \left [ 1+M.\sin \left ( f'.2\pi\frac{i}{f_s -1} \right ) \right] .P .\sin \left ( f.2\pi\frac{i}{f_s -1} \right ) \right \}_0^{\Lambda-1} = \\ 
%                         =  \left\{P.\sin \left( f.2\pi\frac{i}{f_s -1}  \right ) +  \frac{P.M}{2} \left [ \sin \left( (f-f').2\pi\frac{i}{f_s -1}  \right )  + \sin \left( (f+f').2\pi\frac{i}{f_s -1}  \right ) \right ] \right \}_0^{\Lambda-1}
% \end{equation}
% \end{twocolequfloat}%
% \begin{multicols}{2}%
% }

@@

\begin{equation}\label{eq:rhythmicUnit}
P^{ \{ j_k \} }_{ \{ i_{k} \}}
\end{equation}

### 2017-08-01 article/acm/listings.tex
finished listing of equations

>>>>

\begin{table*}[htp!]
\centering
\caption{Equation numbers and their descriptions.
All these equations are implemented in file \texttt{src/sections/2.py}.}
\begin{tabular}{ c | p{12cm} }
   Number & Description \\\hline
 \ref{eq:dur} & relation between number of samples and duration \\
 \ref{eq:potencia} & power of the wave \\
 \ref{decibels} & decibels by difference by means of the power of each wave \\
 \ref{eq:ampVol} & double amplitude implies $\approx 6dB$ \\
 \ref{eq:potVol} & double power implies $\approx 3dB$ \\
 \ref{eq:dobraVol} & double volume ($10dB$) implies a factor of $\approx 3.16$ for amplitude \\
 \ref{ampDec} & direct relation between variations in amplitude and decibels \\
 \ref{periodicidade} & equivalences in a periodic sound with respect to wavelength, frequency and sample rate \\
 \ref{sinusoid} & sample amplitudes in a sinusoid \\
 \ref{sawTooth} & sample amplitudes in a sawtooth wave \\
 \ref{triangular} & sample amplitudes in a triangular wave \\
 \ref{square} & sample amplitude in a square wave \\
 \ref{sampleandoFormaDeOnda} & samples in a sound derived from a sampled wave period \\
 \ref{recomposicaoFourier} & reconstruction of samples from the Fourier components \\
 \ref{moduloEfase} & reconstruction of real samples (e.g. for audio) from Fourier components \\
 \ref{coefsPareados} & number of pairs of Fourier coefficients which are related to the same frequency \\
 \ref{equivalenciasFreqs} & indexes of equivalent frequencies and coefficients for real signals \\
 \ref{equivalenciasModulos} & equal modules between samples of real signals \\
 \ref{equivalenciasFases} & equivalence in phases between samples of real signals \\
 \ref{eq:reconsCompleta} & complete reconstruction of the signal using Fourier components and previous equations \\
 \ref{eq:notaBasica} & sample sequence related to a basic note \\
 \ref{periodoUnico} & sample sequence of a period of an arbitrary waveform \\
 \ref{eq:notaBasicaTimbre} & note samples derived from a sampled waveform \\
 \ref{eq:distOuvidos} & distance of a (sound) source to each ear given the distance between the ears and an (x,y) position of the source \\
 \ref{eq:dti} & Interaural Time Difference (ITD), the time difference of a sound reaching each ear \\
 \ref{eq:dii} & Interaural Intensity Difference (IID), the intensity difference (in decibels) of a sound reaching each ear \\
 \ref{eq:locImpl} & ITD and IID in terms of sample delays and their amplitudes \\
 \ref{eq:angulo} & azimuthal angle of a (x, y) source \\
 \ref{eq:mixagem} & samples that result from mixing sounds \\
 \ref{eq:concatenacao} & samples that result from concatenating sounds \\
\end{tabular}
\end{table*}

\begin{table*}[htp!]
\centering
\caption{Equation numbers and their descriptions.
All these equations are implemented in file \texttt{src/sections/3.py}.}
\begin{tabular}{ c | p{12cm} }
   Number & Description \\\hline
 \ref{eq:lut} & sample sequence generated by means of a lookup table (LUT) \\
 \ref{freqLinear} & frequency at each sample in a linear transition of frequency \\
 \ref{indiceLinear} & indices of a LUT in a linear transition of frequency \\
 \ref{serieAmostralLin} & sample sequence obtained through a LUT in a linear transition of frequency \\
 \ref{freqExponencial} & frequency at each sample in an exponential transition of frequency (linear pitch) \\
 \ref{indiceExponencial} & indices of a LUT in an exponential transition of frequency (linear pitch) \\
 \ref{serieAmostralLog} & sample sequence obtained through a LUT in an exponential transition of frequency \\
 \ref{seqAmp} & amplitude factors at each sample in an exponential transition of amplitude ($\approx$ linear volume) \\
 \ref{transAmp} & sample sequence with an exponential transition of amplitude ($\approx$ linear volume) \\
 \ref{seqAmpLin} & amplitude factors at each sample in a linear transition of amplitude \\
 \ref{seqAmpDB} & sample sequence in an exponential transition of amplitude ($\approx$ linear volume) with difference given in decibels \\
 \ref{eq:conv} & sample sequence obtained through the convolution of two other sequences (e.g. for applying FIR filters) \\
 \ref{eq:diferencas} & difference equation (e.g. for applying IIR filters) \\
 \ref{eq:passa-baixas} & IIR coefficients for a simple, useful and well-behaved low-pass filter \\
 \ref{eq:passa-altas} & IIR coefficients for a simple, useful and well-behaved high-pass filter \\
 \ref{eq:varAux} & auxiliary variables for the following band-pass and band-reject filters \\
 \ref{eq:passa-banda} & IIR coefficients for a simple, useful and well-behaved band-pass filter \\
 \ref{eq:rejeita-banda} & IIR coefficients for a simple, useful and well-behaved band-reject filter \\
 \ref{eq:branco} & Fourier coefficients of a white noise \\
 \ref{eq:rosa} & Fourier coefficients of a pink noise \\
 \ref{eq:marrom} & Fourier coefficients of a brown noise \\
 \ref{eq:azul} & Fourier coefficients of a blue noise \\
 \ref{eq:violeta} & Fourier coefficients of a violet noise \\
 \ref{eq:preto} & Fourier coefficients of a black noise \\
 \ref{vbrGamma} & indices for a vibrato given its frequency and using a LUT \\
 \ref{vbrAux} & samples for applying a vibrato \\
 \ref{vbrF} & frequency at each sample of a sound with vibrato \\
 \ref{vbrGamma2} & indices for LUT in a sound with vibrato \\
 \ref{vbrT} & sample sequence of a sound with vibrato \\
 \ref{trA} & amplitude at each sample for a tremolo \\
 \ref{trT} & sample sequence of a sound with tremolo \\
 \ref{eq:fmEsp} & components in FM synthesis when both modulator and carrier are sines \\
 \ref{eq:Bessel} & the Bessel function \\
 \ref{eq:specAM} & components in AM synthesis when both modulator and carrier are sines \\
 \ref{fmGammaAux} & indices for LUT in modulator of an FM synthesis \\
 \ref{fmAux} & sample sequence of a modulator in an FM synthesis \\
 \ref{fmF} & frequeny at each sample of a sound derived from FM synthesis \\
 \ref{fmGamma} & indices for the final signal in FM synthesis using LUT \\
 \ref{fmT} & sample sequence of a sound generated through FM and using LUT \\
 \ref{amA} & amplitude at each sample in a sound generated though AM \\
 \ref{amT} & sample sequence of a sound generated through AM and using LUT \\
\end{tabular}
\end{table*}

\begin{table*}[htp!]
\centering
\caption{Equation numbers and their descriptions.
All these equations are implemented in file \texttt{src/sections/3.py}.}
\begin{tabular}{ c | p{12cm} }
   Number & Description \\\hline
 \ref{eq:vinculos} & an example of bonds between musical characteristics \\
 \ref{eq:fDoppler} & relation between frequencies and speed in the Doppler effect \\
 \ref{eq:aDoppler} & relation between position, speed and amplitude in the Doppler effect \\
 \ref{eq:ffDoppler} & relation between position, speed and amplitude in the Doppler effect \\
 \ref{eq:p1rev} & samples of a FIR filter for the first period of a reverberation \\
 \ref{eq:p2rev} & samples of a FIR filter for the second period of a reverberation \\
 \ref{eq:rev} & samples of the FIR filter for a reverberation (considering both first and second periods) \\
 \ref{eq:adsr} & amplitude factors for each sample in an ADSR envelope \\
 \ref{eq:adsrApl} & sample sequence of a sound with an ADSR envelope \\
\end{tabular}
\end{table*}

\begin{table*}[htp!]
\centering
\caption{Equation numbers and their descriptions.
All these equations are implemented in file \texttt{src/sections/4.py}.}
\begin{tabular}{ c | p{12cm} }
   Number & Description \\\hline
 \ref{escSim} & perfectly symmetric scales in each octave with the twelve semitones \\
 \ref{eq:escalas} & diatonic scales \\
 \ref{eq:relacaoDia} & the succession of tones and semitones of a diatonic scale \\
 \ref{eq:escalasMenores} & sequences of semitones for the three types minor scales \\
 \ref{eq:serieHarmonica} & harmonic series in terms of semitones \\
 \ref{triades} & triads (chords constituted by thirds) \\
 \ref{eq:rhythmicUnit} & a convention to specify a unit of rhythmic division or agglomeration \\
 \ref{eq:groups} & definition of algebraic groups \\
\end{tabular}
\end{table*}

\subsection{In Section 4 - Notes in music}

\subsection{In Section 3 - Variations of the basic note}
\subsection{In Section 4 - Notes in music}
